\documentclass[12pt]{article}
\usepackage{geometry,amsmath,amssymb, graphicx, natbib, float, enumerate}
\geometry{margin=1in}
\renewcommand{\familydefault}{cmss}
\restylefloat{table}
\restylefloat{figure}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\logit}{\mathrm{logit}}

\begin{document}
\noindent
{\bf BST 140.752 \\ Problem Set 5} \\

\section{LMM}
\begin{enumerate}
\item Let $Y_{ij} = \mu + u_i + \epsilon_{ij}$ for $u_i \sim N(0,\sigma^2_u)$ and $\epsilon_{ij} \sim N(0, \sigma^2)$. Calculate the BLUP 
for $u_i$
\item Let $Y = X \beta + Z u + \epsilon$ for $u \sim N(0,\Sigma_u)$ and $\epsilon \sim N(0, \sigma^2 I)$. Calculate the BLUP for $u$. 
\item Load the Rail data set in R. Fit a mixed model of the form from question 1. Compare a the estimates of the mean for each rail with the empirical mean.	
\item Load the pixel data set in R. Fit a linear mixed effect model where you have $Y_{ijk} = \beta_0 + \beta_1 x_{k} + u_{i} + u_{ij} + \epsilon_{ijk}$
where $Y_{ijk}$ is pixel, $i$ is dog, $j$ is side and $k$ is day index and $x_k$ is day. Fit the model and interpret the results.
\item Consider the model $Y_{i} = \mu + \epsilon_{ij}$. Consider putting a so-called "flat" prior on $\mu$. That is acting like a distribution
that is $1$ from $-\infty$ to $+\infty$ is a valid density. Calculate the distribution marginalized over $\mu$ and show that it is the same likelihood
used to obtain the REML estimates.
\item Derive the BLUP estimates for a linear mixed effect model of the 
form $Y = X\beta + Z U + \epsilon$ where $U \sim N(0, \Sigma_u)$
and $\epsilon \sim N(0, \sigma^2 I)$ and 
assuming known variance components. 
\item Find the BLUPs for the Rail data set in R. Compare the BLUP estimates
with those obtained using a fixed effect estimate for the $U_i$.
\item Consider the model from Question 1. Show that the estimate for 
$\sigma^2$ is not consistent as $i \rightarrow \infty$ for equal
numbers of observations per cluster. 

\end{enumerate}

\section{Smoothing}
\begin{enumerate}
\item Let \texttt{seq(0, 6 * pi, length = 1000)}. Simulate a sin wave and do the
following:
    \begin{enumerate}[A.]
        \item Fit a smoother using linear and quadratic regression splines.
        \item Fit a smoother using bsplines and natural splines.
        \item Plot the columns of the design matrix for the different bases.
    \end{enumerate}
\item Take the \texttt{mtcars} dataset. Fit a penalized spline model using
\texttt{mgcv} with \texttt{mpg} as the outcome and additive smooth functions
of weight and horsepower. Visualize the fits.
\item Show that the function 
$\sum_{k=0}^p \beta_k x^k + \sum_{k=1}^K \gamma_k (x - \xi_k)_+^p$ 
has $p-1$ derivatives for $p>1$. 
\end{enumerate}

\section{Bayesian models}
\begin{enumerate}
\item Let $Y ~|~ \mu, \theta \sim N(\mu J_{n}, \theta^{-1} I)$ be an $n\times 1$ vector. Further
let $\mu ~|~ \theta \sim N(\mu_0 , \theta^{-1} \sigma^2_0)$ and $\theta \sim \mbox{Gamma}(\alpha_0, \beta_0)$.
Derive the joint posterior distribution for $\mu$ and $\theta$ as well as the marginal posterior distribution 
for $\mu$ and $\theta$.
\item Let $p$ denote the unknown proportion of rocks in a riverbed that are
sedimentary in type. Suppose that $X = 12$ of a sample of $n = 20$ rocks collected in
random locations are found to be sedimentary in type. Display the Bayes interval for 
a uniform and Jeffrey's prior. Give 95\% HPD and equi-tail intervals for both.
\item A cohort study was performed to evaluate the relation between oral-contraceptive
use and breast cancer in the age group 40-44. At enrollment, women were classified
as current users, past users or never users of oral contraceptives. Current users had 13 cancer
cases out of 4,671 person years while never users 113 cancer cases out of 98,091 years.
Assume that cancer counts given $t$ person years is Poisson($\lambda t$) where $\lambda$
is the incidence rate. Assume independent Gamma priors for the incident rate for the never
and current users and create the following (trying different values for the hyper-parameters):
\begin{enumerate}[a.]
\item The posterior for the incidence rate for the never users.
\item The posterior for the incidence rate for the current users.
\item The posterior for the relative rate comparing the never to current users.
\item Give 95\% equi-tail credible intervals for all three above.
\end{enumerate}
\item Take the \texttt{mtcars} dataset. Fit a Bayesian regression model with 
\texttt{mpg} as the outcome and \texttt{wt} as the predictor. Display the posterior
for the slope, intercept and residual variance. Assume a normal/gamma prior.
\end{enumerate}


\end{document}

\section{Residuals}
\begin{enumerate}
\item Go through the argument that the propensity score is a balancing score and fill in the details of the proof.
\item Give the number of continuous derivatives for the function $f(x) = \sum_{k=0}^k \beta_k x^k + \sum_{j=1}^J x (x - \xi_j)^k_+$ for known knot points.
\item Generate a sinusoid with noise around it. Use regression splines to fit the model.
\item Consider the model $Y_{ij} = \beta_0 + U_i + \epsilon_{ij}$ where $U_i \sim N(0, \sigma_u^2)$
and $\epsilon_{ij} \sim N(0,\sigma^2$. Derive $E[U_i | Y]$.
\item Consider $Y_i = \mu + \epsilon_i$ where $\epsilon_i \sim N(0, \sigma^2)$. Show that the REML
estimate of $\sigma^2$ is unbiased.
\end{enumerate}

\end{document}
