\documentclass[12pt]{article}
\usepackage{geometry,amsmath,amssymb, graphicx, natbib, float, enumerate}
\geometry{margin=1in}
\renewcommand{\familydefault}{cmss}
\restylefloat{table}
\restylefloat{figure}

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\logit}{\mathrm{logit}}
\newcommand{\RQ}{[{\bf REQUIRED}]~}


\begin{document}
\noindent
{\bf BST 140.752 Midterm exam}  \\ \ \\ \ \\
\hrule
\noindent {\bf printed name}
\\ \ \\
\hrule

\noindent Assumptions
\begin{itemize}
\item $Y = \alpha + X_1 \beta_1 + X_2 \beta_2 + \epsilon = X \beta + \epsilon$ where $X = [J ~ X_1 ~ X_2]$ and $J$ is a vector of ones.
\item $Y$ is $n\times 1$.
\item $X_1$ is a full column rank $n\times p_1$ matrix.
\item $X_2$ is a full column rank $n\times p_2$ matrix.
\item $\epsilon \sim N(0,\sigma^2 I)$.
\item  $X_1' X_2 = 0$ and $X'_1 J = X_2' J = 0$.
\end{itemize}
\ \\ 
\hrule 
\noindent Questions
\begin{enumerate}[1.]
\item Argue that the empirical correlation between any column of $X_1$
  and any column of $X_2$ is 0 and that the mean of the columns of $X_1$ are zero. 
\item Show that
$\hat \alpha$, $\hat \beta_1$ and $\hat \beta_2$ (the least squares estimates) are all independent of one and another and give the estimates and variances.
\item Assume that $\sigma^2$ is known. Derive a Chi-squared test that $\beta_1 = 0$. 
\item An ellipse centered at the point $v \in \mathbb{R}^k$ is defined as the 
solutions $x\in \mathbb{R}^k$ to the equation $$(x - v)' A^{-1} (x - v) = 1$$ 
The eigenvalues of $A$ determine the length of the axes, by the way. 
Use the F-test of the general linear hypothesis $H_0: K\beta_1 = m$ to derive a confidence ellipse for $\beta_1$.
\item Suppose that you had a full rank design matrix $W = [J ~ W_1 ~ W_2]$ where
$W_1$ and $W_2$ are $n\times p_1$ and $n \times p_2$ respectively. Derive linear transformations of $W_1$ and $W_2$ that get you from $W$ to a design
matrix satisfying all of the qualities of $X$. Specifically that $X_1' J = X_2' J = 0$ and $X_1'X_2 = 0$.
\item [BONUS] Give a setting where, by virtue of the design of an experiment, that $X_1$ would be exactly or nearly uncorrelated with $X_2$.
\end{enumerate}
\hrule

%\newpage



\end{document}

